---
title: "Lab 10 - Point and Interval Estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab 10 - Maximum Likelihood Estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fig_width: 6 
fig_height: 4
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4,
  fig.align = "center"
)
```

```{r setup}
library(MATH4753ROSAlab10)
```

# Introduction

This package produces point and interval estimates from a variety of different functions. Point and interval estimates are a very important part of classical and Bayesian statistics. In this vignette we will investigate different methods of making point and interval estimates according to Maximum Likelihood Theory. That is:

  a. Grid Approximation
  b. Newton-Raphson Method
  c. Algebraic First Principals
  
Though, we will forego the latter in this package. Throughout this vignette, we will utilize various log-likelihood functions as "support" or "helper" functions to the heavy-weight functions that this package is most concerned with.

# `mymaxlik`

We will look at two examples of this function using different distributions: Binomial and Poisson.

## Binomial

```{r}
logbin <- function(x,param) log(dbinom(x, prob = param, size = 20))

v <- mymaxlik(x=c(9,10,11),
         param=seq(0,1,length=1000),
         lfun=logbin,
         xlab=expression(pi),
         main="Binomial",
         cex.main=1)

v
```

## Poisson

```{r}
logpoiss <- function(x,param) log(dpois(x,lambda = param))

k <- mymaxlik(x = c(5,5,6,7),
              param=seq(0,100,length=1000),
              lfun=logpoiss,
              xlab=expression(lambda),
              main="Poisson",
              cex.main=2
              )
k
```

For both functions, the returned list:

* `i` is the index for the value of $x$ which maximizes the function parameter, $p$.

* `parami`, at the end, becomes the value of the parameter which we seek

* `yi` is a vector of column [enter appropriate function here], and $y_i$ is the value in the "`ith`" location of the vector.

* `slope` is a vector of slope values, note it contains three positive values and two non-positive ones.

# `mymld`

```{r}
logpoiss = function(x,param) log(dpois(x,lambda=param)) 

l <- mymld(x = c(5,5,6,7),
           param=seq(0,100,length=1000),
           lfun=logpoiss
           )
l
```

This is a more accurate version of the previous previous function. The returned value indicates the value of the desired parameter (label $\theta$ here) which _approximately_ gives the maximum.

# `mymaxlikg`

```{r}
logbin2=function(theta){log(dbinom(4,prob=theta,size=10)) + log(dbinom(12,prob=theta,size=20)) + log(dbinom(7, prob = theta, size = 15))}

j <- mymaxlikg(theta=seq(0,1,length=10000))
j
```








